---
title: "THE SPARKS FOUNDATION - GripJune21"
author: "TASK 1 - Prediction using Supervised Machine Learning"
output:
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Predict the percentage of a student based on the no. of study hours.

#### Done By: Jerin M Steefan

### **Importing the Packages**
```{r}
library(dplyr)
library(ggplot2)
library(corrplot)
```


### **Importing the Dataset**
```{r}
# loading the dataset from the url provided

data = read.csv(url("http://bit.ly/w-data"))
head(data)
```

```{r}
#taking the dimension of the dataset

dim(data)
```
 *The dataset contains 25 observations with 2 variables.*
 
```{r}
#finding the Column names

colnames(data)
```
*Hours and Scores are the two variables present in the dataset.*


### **DATA PREPROCESSING**

```{r}
#checking for NAN values

colSums(is.na(data))
```

*The given dataset contains no NaN values.*

```{r}
#checking NULL values

is.null(data)
```
*The given dataset contain no NULL values.*

### **EXPLORATORY DATA ANALYSIS**

```{r}
#structure of the dataset

str(data)
```
*The variable Hours is of type Numeric and Scores is of datatype Integer.*


```{r}
#SUMMARY OF THE DATASET

summary(data)
```
  * The Minimum value of Hours is 1.100 and maximum value is 9.200. Mean value is greater than median value. Hence it      is right skewed.
  * The Minimum value of Scores is 17.00 and maximum value is 95.00. Mean value is greater than median value. Hence it      is right skewed.

**BOXPLOT**
```{r}
# Checking the outliers

library(ggplot2)
boxplot(data,main='Boxplot of Hours and Scores')
```

*From the boxplot, we can understand that No outliers are present in the dataset.*


**SCATTERPLOT**
```{r}

library(ggplot2)
my_graph <- ggplot(data,
                    aes(x =Hours, y =Scores))+
                    ggtitle('HOURS V/S SCORES')+
                    theme(plot.title = element_text(hjust = 0.5))+
                    geom_point(col='red') 
  
my_graph
```

*From the graph we can see that there is a linear relationship between the response variable and explanatory variable. Also the direction of association seems to be positive i.e. As Hours increase, the Scores obtained also increase and vice-versa.*


**CORRELATION**
```{r}
library(corrplot)
corrplot(cor(data),
method ='number',
type = 'lower' # show only lower side
)
```

```{r}
#using pearson correlation

cor.test(data$Hours,data$Scores)
```
*Here, the correlation value is 0.9761907. Hence, we can understand that there exists a high positive correlation between Hours and Scores.*

### **DATA MODELLING**

*Train Test Splitting*
```{r}
set.seed(100)
rows=sample(nrow(data))

#Randomly order data
data=data[rows,]
#Identify row to split on: split
split = round(nrow(data) * .80)
```

```{r}
#Create train
train=data[1:split,]
#Create test
test=data[(split+1):nrow(data),]

```


**Linear Regression Model**
```{r}
#fitting linear regression model
linmod = lm(Scores~Hours, data = train)

#taking the summary of the model
summary(linmod)
```
The value of intercept of the linear model is 3.5030. The slope of the model is 9.4682.
Hence,the model can be interpreted as :
 $Scores = 9.4682 * Hours + 3.5030$. 

  * Residual standard error is the measure of the quality of the linear regreesion fit and here it is $5.508$ on $18$      degrees of freedom.
  * R squared statistic provides measure of how well the model is fitting the actual dataset.Here, $94%$ of fitting to     the linear model.
  * F statistic value is $311.3$, which is relatively larger than 1. Hence, a good relationship is existing between        Sales and Spend.


*Predicting the Scores*
```{r}
Pred = predict(linmod, test)
```


*Comparing The Actual and Predicted Scores*
```{r}
data.frame(Actual=test$Scores,Predicted=Pred)
```

*Comparing The Actual and Predicted Values using Data Visualisation*
```{r}
library(ggplot2)
my_graph <- ggplot(train,
                    aes(x =Hours, y =Scores))+
                    ggtitle('Actual V/S Predicted')+
                    theme(plot.title = element_text(hjust = 0.5))+
                    geom_point(col='red')+
                    stat_smooth(method = 'lm',
                    col = 'blue',
                    se = FALSE,
                    size = 1)
  
my_graph
```

#### **What will be the predicted score if the student studies for 9.25 hr/day?**
```{r}
test2 = data.frame(Hours = 9.25)
predict(linmod, test2)
```
*Therefore, according to the regression model, if a student studies for 9.25 hours per day he/she is likely to score 91.08419.*



